# Reproducible Research: Peer Assessment 1

## Set echo=TRUE for all chuncks to show R codes in markdown
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r set_directory, echo = FALSE}
setwd("~/Desktop/coursera/JHU_R/reproducible_research/week2/assignment/RepData_PeerAssessment1")
```

## Loading and preprocessing the data
```{r get and clean data}
data <- read.csv("activity.csv")
data$date <- as.Date(data$date,"%Y-%m-%d")
```


## What is mean total number of steps taken per day?
```{r total_daily_steps, fig.height = 4}
daily_step <- tapply(data$steps, data$date, sum, na.rm = TRUE)
hist(daily_step,breaks = 10, col = "lightblue", main = "histogram of total steps per day", xlab = "number of steps per day")
daily_step_mean <- mean(daily_step)
daily_step_median <- median(daily_step)
```

The mean total daily steps is `r daily_step_mean`, and the median of total daily steps is `r daily_step_median`.

## What is the average daily activity pattern?
```{r daily_pattern, fig.height = 4}
mean_inter_step <- tapply(data$steps, data$interval, mean, na.rm = TRUE)
plot(unique(data$interval), mean_inter_step, type = "l", col="blue", main = "average daily activity pattern",
     xlab = "interval", ylab = "average steps")
max_interval <- data$interval[which.max(mean_inter_step)]
```

The `r max_interval`th contains the maximum number of steps across all days.

## Imputing missing values
```{r imputing_NA, fig.height = 4}
#fill NA with mean of the interval
num_NA <- summary(data)[7,1]
library(plyr)
data2 <- ddply(data, .(interval), function(df){
    df$steps[is.na(df$steps)] <- mean(df$steps, na.rm=TRUE)
    return (df)
})
daily_step2 <- tapply(data2$steps, data2$date, sum, na.rm = TRUE)
hist(daily_step2,breaks = 10, col = "lightblue", main = "histogram of total steps per day (NA filled)", xlab = "number of steps per day")
daily_step2_mean <- mean(daily_step2)
daily_step2_median <- median(daily_step2)
```

The total number of "NA" is `r num_NA`.By filling in the "NA"" values, the new mean total daily steps is `r daily_step2_mean`, and the new median of total daily steps is `r daily_step2_median`, both larger than the original "NA" removed results.

## Are there differences in activity patterns between weekdays and weekends?
```{r weekd_comparison, fig.height = 6}
data2$days <- factor(c("weekend","weekday"))
library(chron)
data2$days[is.weekend(data2$date)] <- as.factor("weekend")
data2$days[!is.weekend(data2$date)] <- as.factor("weekday")
data3 <- ddply(data2, .(days, interval),summarize, mean_step = mean(steps))
library(ggplot2)
p <- ggplot(data3, aes(x = interval, y = mean_step))
p + geom_line(colour = "blue") + facet_grid(days ~.) +labs(title="daily activity patterns",y="number of steps")
```

There are some differences in activity patterns between weekdays and weekends.
